---
title: "Anthropometrics Changes From Baseline"
subtitle: Program 09
author: "Francois Collin"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
---

## Introduction

Diabetes group estimations in anthropometrics changes from baseline were
obtained and tested by analysis of covariance models (Ancova), relying once
more on the Fisher test; to increase accuracy and statistical power,
estimations were adjusted for baseline values
[@Vickers2001; @VanBreukelen2006; @CHMP2015; @OConnell2017].

The analysis was realized with R [v4.2.0, @RCT2022]. Note the
use of the additional packages downloaded from the RStudio package manager
repository (freeze date 2022-05-04, [repos](
https://packagemanager.rstudio.com/all/2022-05-04+Y3JhbiwyOjQ1MjYyMTU7NTVFQjhGODU)):

- `car` for regression model test [version 3.0-13, @Fox2019]
- `emmeans` for least-Squares Means (LSM) estimations [version 1.7.3, @Lenth2022]
- `flextable` for output table formatting [version 0.7.0, @Gohel2022]
- `tidyr` for data wrangling [version 1.2.0, @Wickham2022].

## Settings

```{r, settings}
cfg_prog <- yaml::read_yaml("_prog.yml")
devtools::load_all("src/pkg/dbs.data")
devtools::load_all("src/pkg/latarnia.utils")
source("R/inches.R")

knitr::opts_chunk$set(results = cfg_prog$knitr$results)
```

## Data

```{r, adam}
#| code-fold: show
adsl <- dbs.data::adsl
advs <- dbs.data::advs
adlb <- dbs.data::adlb
```

```{r, ads}
ads <- adlb |>
  rbind(advs) |>
  subset(basetype == "" & avisit != "Baseline") |>
  subset(select = c(subjid, paramcd, avisit, base, chg)) |>
  (\(x) merge(x = adsl[c("subjid", "diabcd")], y = x, by = "subjid"))() |>
  (\(df, fct = "diabcd") {
    df[paste0(fct, "_n")] <- factor_n(df, fct, id = "subjid", sep = " ")
    df
  })()

head(ads)
```

## Helper functions

```{r, extra_functions}
format_pval <- function(x) {
  p <- round(x, 5)
  ifelse(
    test = p < 0.0001,
    yes = "<0.0001",
    no = ifelse(
      test = p < 0.001,
      yes = format(round(p, 4), nsmall = 4),
      no = ifelse(
        test = p < 0.01,
        yes = format(round(p, 3), nsmall = 3),
        no = format(round(p, 2), nsmall = 2)
      )
    )
  )
}

lm_by_paramcd <- function(x,
                          dep_var = "aval",
                          indep_var = "diabcd_n",
                          covariate = NULL) {
  
  formula <- paste(
    dep_var, "~", 
    if (!is.null(covariate)) paste(covariate, "+"),
    indep_var
  )
  formula <- as.formula(formula)
  
  lapply(x, \(x) list(data = x, lm = lm(formula, data = x)))
}

make_specs <- function(var) as.formula(paste("~", var))
lsm_by_param <- function(x, indep_var = "diabcd_n") {
  lapply(
    x,
    \(x) {
      mod_em <- emmeans::emmeans(x$lm, specs = make_specs(indep_var))
      y <- as.data.frame(mod_em)
      cbind(
        paramcd = unique(x$data$paramcd),
        y,
        diabcd_f = car::Anova(x$lm)[indep_var, "Pr(>F)"]
      )
    }
  )
}
```

## Tab 09 01 - Ancova - Anthropometrics Changes from Baseline by Diabetes Group

```{r}
tab_09_01_raw <- ads |>
  (\(x) split(x, f = x$paramcd))() |>
  lm_by_paramcd(dep_var = "chg", covariate = "base", indep_var = "diabcd_n") |>
  lsm_by_param(indep_var = "diabcd_n") |>
  (\(x) Reduce(rbind, x))()

tab_09_01 <- tab_09_01_raw |>
  (\(df) df[order(df$diabcd_n), ])() |>
  within({
    val <- paste0(
      signif(emmean, 3),
      " (", signif(lower.CL, 3), ", ", signif(upper.CL, 3), ")"
    )
    pval <- format_pval(diabcd_f)
  }) |>
  tidyr::pivot_wider(
    id_cols = c("paramcd", "pval", "diabcd_f"),
    values_from = "val",
    names_from = "diabcd_n"
  ) |>
  (\(x) x[order(x$diabcd_f), ])()
tab_09_01
```

```{r}
library(flextable)
wrap_line <- function(x) paste(strwrap(x, width = 80), collapse = " ")
tab_09_01_ft <- tab_09_01 |>
  subset(select = -diabcd_f) |>
  flextable() |>
  autofit() |>
  footnote(
    part = "header",
    i = 1, j = 2,
    value = as_paragraph(
      "Note: pval, p value of diabetes group effect test by F test."
    ),
    ref_symbols = "a"
  ) |>
  add_footer_lines(c(
    wrap_line("Source: Full Analysis Set, observed cases at baseline and post
    intervention."),
    "Note: rows are ordered by increasing p values, most significant on top."
  )) |>
  add_header_lines("Analysis Set: Full Analysis Set - Observed Cases") |>
  set_caption(
    caption =  wrap_long_lines(
      "Tab 09 01 - Analysis of Covariance / Least Means Square estimations of
      Anthropometrics Changes from Baseline by Diagnosis Group at
      Month 3 (95% Confidence Interval) Adjusted for Baseline"
    )
  ) |>
  theme_booktabs()
tab_09_01_ft
```

```{r}
bnm <- "tab_09_01"
dir_tab <- cfg_prog$paths$tab
dir_dta <- cfg_prog$paths$dta

file.path(dir_tab, paste0(bnm, "_ft.RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tab_09_01_ft, file = .)

file.path(dir_dta, paste0(bnm, ".RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tab_09_01, file = .)

file.path(dir_tab, paste(bnm, sep = ".", "docx")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_docx(tab_09_01_ft, path = .)

file.path(dir_tab, paste(bnm, sep = ".", "html")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_html(tab_09_01_ft, path = .)

file.path(dir_dta, paste(bnm, sep = ".", "csv")) %T>%
  message("[output] Table saved as ", .) %>%
  write.csv(tab_09_01, file = ., row.names = FALSE)
```

```{r}
detach(package:flextable)
```


## Example: Glucose 120 - Fig 09 01

```{r, f09_01}
library(ggplot2)
dta <- dbs.data::adlb |>
  subset(
    paramcd == "GLU120" &
      dtype == "" &
      avisit == "Month 3" &
      basetype == ""
  ) |>
  merge(x = adsl[c("subjid", "diabcd")], y = _, by = "subjid")

dta$lm_pred <- predict(lm(aval ~ base, data = dta))
lim <- range(c(dta$aval, dta$base))

gg1 <- ggplot(dta, aes(base, aval, color = diabcd)) +
  geom_point() +
  ylab("Post exercise intervention") +
  xlab("Baseline") +
  geom_segment(aes(xend = base, yend = lm_pred)) + 
  geom_smooth(method = "lm", se = FALSE, color = "black", linetype = 1) +
  scale_color_viridis_d(begin = .1, end = .9, direction = -1) +
  coord_cartesian(xlim = lim, ylim = lim) +
  theme_minimal() +
  theme(legend.position = "top")

gg2 <- ggplot(dta, aes(base, fill = diabcd, color = diabcd)) +
  geom_boxplot(alpha = .5, show.legend = FALSE) +
  coord_cartesian(xlim = lim) +
  ylab("diabcd") +
  scale_color_viridis_d(begin = .1, end = .9, direction = -1) +
  scale_fill_viridis_d(begin = .1, end = .9, direction = -1) +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.title.x = element_blank(),
    axis.text.y = element_blank(),
    panel.grid = element_blank()
  )

gg <- cowplot::plot_grid(
  gg1, gg2,
  nrow = 2,
  align = "v",
  rel_heights = c(1, .15)
)
p <- clean_slate() |>
  add_header(c("FCA Collin", "UMB BESD"), c("Confidential", "Draft")) |>
  add_title(
    c(
      "Figure 9.1",
      strwrap(
        "Scatter plot - Glucose 120 at Month 3 in relation to Baseline by
        treatment group", width = 80
      ),
      "Analysis Set: Full Analysis Set - Observed Cases"
    )
  ) |>
  add_figure(gg, height = inches(5.5), width = inches(4.5))|>
  add_note(c(
    "Note: the table prog 08 output 01 focuses on the x-axis and check for
     differences at baseline, the boxplots gives an overview of these
     baseline variation.",
    "Note: the table tab prog 09 output 01 shows the effect of exercise in
     each group. The p.values is provided and accounts for the diagnostic group
     effect, while taking into account the fact that the higher the value at
     baseline, the higher after the exercise intervention. Actually, 
     the analysis of covariance model focus on how much above or below the
     general regression (black line) is positionned a diagnosis group.
     E.g., +20 for T2D indicates that the result after exercise is 20 units
     higher than expected if there were no diagnostic group effect; ... however,
     the confidence interval extend from -13.5 and +53.8 which means that
     i) if we repeated the study, 100 times, it would result 95 times from
     -13.5 to 53.8
     ii) the interval includes 0 so we can't conclude to a significant
     increase or decrease in Glucose 120 due to the exercise intervention."
  )) |>
  add_footer(
    c(
      "Program t2d_09_chg / Env ayup_dbs:v0.1.0-alpha",
      format(Sys.time(), format = "%Y-%m-%d %H:%M (%Z)")
    ),
    cfg_prog$version
  )

export_as(
  p,
  file = file.path(cfg_prog$paths$grh, "fig_09_01.pdf"),
  file_graph_alone = file.path(cfg_prog$paths$grh, "fig_09_01_af.pdf")
)

```

```{r, f09_01_sh, eval=cfg_prog$knitr$results!="hide"}
#| fig-width: 8.3
#| fig-height: 11.7
show_slate(p)
```

## Session Informations

```{r}
sessioninfo::session_info()
```

## Untidied information about analysis of change from baseline

_Questions:
for the analysis of a post-treatment values, should we analyse the change
from baseline or percentage change from baseline? Should we adjust for the
baseline?_

Back to 2003, in the context of randomized clinical trial, 
the European Medicines Agency [@CPMP2003],
when the endpoint is studied as a change from
baseline, the adjustment for baseline improves the accuracy in comparison
to non-baseline adjustment; estimates becomes also equivalent to the standard
linear model, the choice of change from baseline analysis or raw value is then
only a question of interpretability. 
They renewed the recommendation in 2015 [@CHMP2015].

From the academic side, the topic was repeatedly studied:

- @VanBreukelen2006:
> _"In randomized studies both methods [Anova, no BL adjustment vs Ancova]
> are unbiased, but ANCOVA has more power"_

- @Liu2009 also highlighted the benefits of adjustment for baseline as a
covariate.
- this was later confirmed by @Zhang2014.
- More recently @OConnell2017 also defended the superiority of the
Ancova-change:
$y_i = \beta_0 + \beta_1 X_i + \beta_2 Y_{0i,=BL} + \varepsilon_i$

> _"Consistent with existing literature, our results demonstrate that each
> method leads to unbiased treatment effect estimates, and based on precision
> of estimates, 95% coverage probability, and power, ANCOVA modeling of either
> change scores or post-treatment score as the outcome, prove to be the most
> effective._"

Most of the authors above are specifically working on randomized trial,
@Vickers2001 also brought some light on the topic, and highlighted in addition
that: working with percentage change is generally a bad idea. The extended
to a theoretical works also indicated that the percentage change from
baseline _"will also fail to protect from bias in the case of baseline imbalance
and will lead to an excess of trials with non-normally distributed outcome
data"_.
