---
title: "Missingness in adlb and advs"
subtitle: Program 03
author: "Francois Collin"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
---

## Introduction

Data missingness was addressed by a missing‑data imputation algorithm,
employed to impute missing values while minimizing bias on results
[R package `missMDA`, @Josse2016]);
"observed cases" and "imputed data" later refers to the exclusion/inclusion
of imputed data.

The analysis was realized with R [v4.2.0, @RCT2022]. Note the
use of the additional package downloaded from the RStudio package manager
repository (freeze date 2022-05-04, [repos](
https://packagemanager.rstudio.com/all/2022-05-04+Y3JhbiwyOjQ1MjYyMTU7NTVFQjhGODU)):

- `ggplot2` for graphics [version 3.3.6, @Wickham2016].

## Program settings

```{r, settings}
params <- yaml::read_yaml("_prog.yml")
devtools::load_all("src/pkg/dbs.data")
devtools::load_all("src/pkg/latarnia.utils")

knitr::opts_chunk$set(results = params$knitr$results)
```

## Data Preparation

```{r, ads}
#| code-fold: show
adsl <- dbs.data::adsl
advs <- dbs.data::advs
adlb <- dbs.data::adlb
```


## Figure 03 output 01

```{r, fig_03_01}
library(ggplot2)
gg <- rbind(adlb, advs) |>
  ggplot(aes(subjid, paramcd, fill = dtype)) +
  scale_fill_manual(values = c("white", "gray75")) +
  geom_tile(color = "gray50") +
  theme_minimal() +
  theme(
    text = element_text(size = 7),
    axis.title = element_blank(),
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.position = "none",
    legend.title = element_blank()
  )
```


```{r, fig_03_01_an}
p <- clean_slate() |>
  add_header(c("FCA Collin", "UMB BESD"), c("Confidential", "Draft")) |>
  add_title(
    c(
      "Figure 3.1",
      "Colored raster - Data missingness per Subject and Parameter",
      "Analysis Set: Full Analysis Set"
    )
  ) |>
  add_figure(gg, width = unit(5.2, "inches"), height = unit(3.5, "inches")) |>
  add_footer(
    paste0("Program t2d_03_bds / Env ayup_dbs:", params$dock$version),
    params$version
  )
```

```{r, fig_03_01_xp}
export_as(
  p,
  file = file.path(params$paths$grh, "fig_03_01.pdf"),
  file_graph_alone = file.path(params$paths$grh, "fig_03_01_af.pdf")
)
```

```{r, fig_03_01_sh, eval=params$knitr$results!="hide"}
#| fig-width: 8.3
#| fig-height: 11.7
show_slate(p)
```

## Session Informations

```{r}
sessioninfo::session_info()
```
