---
title: "Demographics and Baseline Anthropometrics"
subtitle: Program 08
author: "Francois Collin"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
---

## Demographics and Baseline Anthropometrics

Demographics data characterized the diabetes groups in terms of age, number of
training composing the exercise intervention, BMI and diet at baseline. This
was completed by a fine description of the baseline anthropometrics leading
differences between diabetes groups. Both characterizations relied on one-way
analysis of variance, the diabetes effect significance was ruled by a Fisher
test, least mean square estimations were obtained for every diabetes group
along with their 95% confidence interval, and pairwise difference estimation
and significance relied on Tukey’s method.

The analysis was realized with R [v4.2.0, @RCT2022]. Note the
use of the additional packages downloaded from the RStudio package manager
repository (freeze date 2022-05-04, [repos](
https://packagemanager.rstudio.com/all/2022-05-04+Y3JhbiwyOjQ1MjYyMTU7NTVFQjhGODU)):

- `car` for regression model test [version 3.0-13, @Fox2019]
- `emmeans` for least-Squares Means (LSM) estimations [version 1.7.3, @Lenth2022]
- `flextable` for output table formatting [version 0.7.0, @Gohel2022]
- `tidyr` for data wrangling [version 1.2.0, @Wickham2022].


Target:

- [x] Table: Demographics and baseline anthropometrics are tested via an Anova.
- [x] Supp. Table: Post-hoc estimations / tests by diabetes groups.
- [x] Supp. Table: extension of the anova to additional ADVS/ADLB parameters.

Specifications:

- Variable [order](https://docs.google.com/document/d/1zTa6J7NO-Wt_wx5Wqq574E_VcCt7lg-75Nt5BqbLApw/edit?usp=sharing)


```{r, settings}
params <- yaml::read_yaml("_prog.yml")
devtools::load_all("src/pkg/dbs.data")
devtools::load_all("src/pkg/latarnia.utils")

knitr::opts_chunk$set(results = params$knitr$results)

```


```{r, adam}
#| code-fold: show
adsl <- dbs.data::adsl
advs <- dbs.data::advs
adlb <- dbs.data::adlb
```


```{r, ads}
ads <- adlb |>
  rbind(advs) |>
  subset(dtype == "" & avisit == "Baseline") |>
  subset(select = c(subjid, paramcd, avisit, aval, dtype)) |>
  rbind(
    adsl |>
      subset(select = -c(diab, diabcd)) |>
      tidyr::pivot_longer(
        cols = c("age", "trainn"), names_to = "paramcd", values_to = "aval"
      ) |>
      within(dtype <- "") |>
      within(avisit <- "Baseline")
  ) |>
  (\(x) merge(x = adsl[c("subjid", "diabcd")], y = x, by = "subjid"))() |>
  (\(df, fct = "diabcd") {
    df[paste0(fct, "_n")] <- factor_n(df, fct, id = "subjid", sep = " ")
    df
  })() |>
  within(
    paramcd <- factor(
      paramcd,
      levels = c(
        "age", "trainn", "GLU0", "GLU30", "GLU60", "GLU120",
        "INSULIN0", "INSULIN30", "INSULIN60", "INSULIN120",
        "HBA1C", "HOMAB", "HOMAIR", "MATSUDA",
        "TRIG", "CHOL", "LDL", "HDL", "VO2MAXE", "WEIGHT", "BMI",
        "FATMASS", "BODYFATP", "LEANMASS", "VAT",
        "DCAL", "FATACFR", "DCARBT", "DFATT", "DPROT", "SMMASS",
        "VO2MAXLBM", "VO2MAXML"
      )
    ) 
  ) |>
  (\(df) df[order(df$paramcd), ])()

```


```{r, extra_functions}
format_pval <- function(x) {
  p <- round(x, 5)
  ifelse(
    test = p < 0.0001,
    yes = "<0.0001",
    no = ifelse(
      test = p < 0.001,
      yes = format(round(p, 4), nsmall = 4),
      no = ifelse(
        test = p < 0.01,
        yes = format(round(p, 3), nsmall = 3),
        no = format(round(p, 2), nsmall = 2)
      )
    )
  )
}

lm_by_paramcd <- function(x,
                          dep_var = "aval",
                          indep_var = "diabcd_n",
                          covariate = NULL) {
  
  formula <- paste(
    dep_var, "~", 
    if (!is.null(covariate)) paste(covariate, "+"),
    indep_var
  )
  formula <- as.formula(formula)
  
  lapply(
    x,
    \(x) list(data = x, lm = lm(formula, data = x))
  )
}

make_specs <- function(var) as.formula(paste("~", var))
lsm_by_param <- function(x, indep_var = "diabcd_n") {
  lapply(
    x,
    \(x) {
      mod_em <- emmeans::emmeans(x$lm, specs = make_specs(indep_var))
      y <- as.data.frame(mod_em)
      cbind(
        paramcd = unique(x$data$paramcd),
        y,
        diabcd_f = car::Anova(x$lm)[indep_var, "Pr(>F)"]
      )
    }
  )
}

lsm_pairs_by_param <- function(x, indep_var = "diabcd_n")
  lapply(
    x,
    \(x) {
      mod_em <- emmeans::emmeans(
        x$lm, specs = indep_var, contr = "revpairwise"
      )
      y <- merge(
        as.data.frame(mod_em$contrast)[c("contrast", "p.value")],
        confint(mod_em)$contrasts
      )
      cbind(paramcd = unique(x$data$paramcd), y)
    }
  )
```


## Tab 08 01 - Demographics and Baseline Anthropometrics by Diabetes Group

```{r}
tab_08_01_raw <- ads |>
  subset(
    paramcd %in% c(
      "age", "trainn", "GLU0", "GLU30", "GLU60", "GLU120",
      "INSULIN0", "INSULIN30", "INSULIN60", "INSULIN120",
      "HBA1C", "HOMAB", "HOMAIR", "MATSUDA",
      "TRIG", "CHOL", "LDL", "HDL", "VO2MAXE", "WEIGHT", "BMI",
      "FATMASS", "BODYFATP", "LEANMASS", "VAT"
    )
  ) |>
  (\(x) split(x, f = x$paramcd, drop = TRUE))() |>
  lm_by_paramcd() |>
  lsm_by_param() |>
  (\(x) Reduce(rbind, x))()

library(tidyr)
tab_08_01 <- tab_08_01_raw |>
  (\(df) df[order(df$diabcd_n), ])() |>
  within({
    val <- paste0(
      signif(emmean, 3),
      " (", signif(lower.CL, 3), ", ", signif(upper.CL, 3), ")"
    )
    pval <- format_pval(diabcd_f)
  }) |>
  pivot_wider(
    id_cols = c("paramcd", "pval", "diabcd_f"),
    values_from = "val",
    names_from = "diabcd_n"
  )
tab_08_01
```

```{r}
library(flextable)
tab_08_01_ft <- tab_08_01 |>
  subset(select = -diabcd_f) |>
  flextable() |>
  autofit() |>
  add_header_lines(wrap_long_lines(
    "Analysis Set: Full Analysis Set - Observed Cases at baseline"
  )) |>
  set_caption(
    caption = wrap_long_lines(
      "Tab 08 01 - Analysis of Variance / Least Means Square estimations
      (95% Confidence Interval) of Demographics Parameters and Baseline
      Anthropometrics by Diabetes Group"
    )
  ) |>
  footnote(
    part = "header",
    i = 2, j = 2,
    value = as_paragraph(
      "Note: pval, p value of diabetes group effect test by F test."
    ),
    ref_symbols = "a"
  ) |>
  footnote(
    value = as_paragraph(
      "Source: ADSL and ADVS/ADLB observed cases at baseline."
    ),
    ref_symbols = ""
  ) |>
  theme_booktabs()
tab_08_01_ft
```

```{r}
bnm <- "tab_08_01"
dir_tab <- params$paths$tab
dir_dta <- params$paths$dta

file.path(dir_tab, paste0(bnm, "_ft.RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tab_08_01_ft, file = .)

file.path(dir_dta, paste0(bnm, ".RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tab_08_01, file = .)

file.path(dir_tab, paste(bnm, sep = ".", "docx")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_docx(tab_08_01_ft, path = .)

file.path(dir_tab, paste(bnm, sep = ".", "html")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_html(tab_08_01_ft, path = .)

file.path(dir_dta, paste(bnm, sep = ".", "csv")) %T>%
  message("[output] Table saved as ", .) %>%
  write.csv(tab_08_01, file = ., row.names = FALSE)
```


## Tab 08 02 - Post-hoc: Demographics and Baseline Anthropometrics by Diabetes Group

```{r}
tab_08_02_raw <- ads |>
  subset(
    paramcd %in% c(
      "age", "trainn", "GLU0", "GLU30", "GLU60", "GLU120",
      "INSULIN0", "INSULIN30", "INSULIN60", "INSULIN120",
      "HBA1C", "HOMAB", "HOMAIR", "MATSUDA",
      "TRIG", "CHOL", "LDL", "HDL", "VO2MAXE", "WEIGHT", "BMI",
      "FATMASS", "BODYFATP", "LEANMASS", "VAT"
    )
  ) |>
  (\(x) split(x, f = x$paramcd, drop = TRUE))() |>
  lm_by_paramcd(indep_var = "diabcd") |>
  lsm_pairs_by_param(indep_var = "diabcd")  |>
  (\(x) Reduce(rbind, x))()

library(tidyr)
tab_08_02 <- tab_08_02_raw |>
  subset(select = c(
    paramcd, contrast, estimate, SE, df, p.value, lower.CL, upper.CL
  ))

library(flextable)
tab_08_02_ft <- tab_08_02 |>
  (\(x) split(x, f = x$paramcd))() |>
  lapply(
    \(x) {
      x$p.value <- format(round(x$p.value, 5))
      x$estimate <- format(signif(x$estimate, 5))
      x$SE <- format(signif(x$SE, 6))
      x$lower.CL <- format(signif(x$lower.CL, 5))
      x$upper.CL <- format(signif(x$upper.CL, 5))
      x
    }) |>
  (\(x) Reduce(rbind, x))()|>
  flextable() |>
  fontsize(size = 9, part = "all") |>
  autofit() |>
  add_header_lines(
    "Analysis Set: Full Analysis Set - Observed Cases at baseline"
  ) |>
  set_caption(
    caption = wrap_long_lines(
      "Tab 08 02 - Post-hoc tests for the Analysis of Variance of
      Demographics and Baseline Anthropometrics by Diabetes Group"
    )
  ) |>
  add_footer_lines(c(
    "CL, 95% Confidence Limit; SE, Standard Error.",
    wrap_long_lines(
      "Note: P value adjustment by Tukey's method for comparing a family of
      3 estimates."
    ),
    "Source: ADSL and ADVS/ADLB observed cases at baseline." 
  )) |>
  theme_booktabs()

tab_08_02_ft
```

```{r}
bnm <- "tab_08_02"
dir_tab <- params$paths$tab
dir_dta <- params$paths$dta

file.path(dir_tab, paste0(bnm, "_ft.RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tab_08_02_ft, file = .)

file.path(dir_dta, paste0(bnm, ".RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tab_08_02, file = .)

file.path(dir_tab, paste(bnm, sep = ".", "docx")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_docx(tab_08_02_ft, path = .)

file.path(dir_tab, paste(bnm, sep = ".", "html")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_html(tab_08_02_ft, path = .)

file.path(dir_dta, paste(bnm, sep = ".", "csv")) %T>%
  message("[output] Table saved as ", .) %>%
  write.csv(tab_08_02, file = ., row.names = FALSE)
```


## Tab 08 03 - Demographics and Baseline Anthropometrics by Diabetes Group (additional parameters)

```{r}
tab_08_03_raw <- ads |>
  subset(
    !paramcd %in% c(
      "age", "trainn", "GLU0", "GLU30", "GLU60", "GLU120",
      "INSULIN0", "INSULIN30", "INSULIN60", "INSULIN120",
      "HBA1C", "HOMAB", "HOMAIR", "MATSUDA",
      "TRIG", "CHOL", "LDL", "HDL", "VO2MAXE", "WEIGHT", "BMI",
      "FATMASS", "BODYFATP", "LEANMASS", "VAT"
    )
  ) |>
  (\(x) split(x, f = x$paramcd, drop = TRUE))() |>
  lm_by_paramcd() |>
  lsm_by_param() |>
  (\(x) Reduce(rbind, x))()

library(tidyr)
tab_08_03 <- tab_08_03_raw |>
  (\(df) df[order(df$diabcd_n), ])() |>
  within({
    val <- paste0(
      signif(emmean, 3),
      " (", signif(lower.CL, 3), ", ", signif(upper.CL, 3), ")"
    )
    pval <- format_pval(diabcd_f)
  }) |>
  pivot_wider(
    id_cols = c("paramcd", "pval", "diabcd_f"),
    values_from = "val",
    names_from = "diabcd_n"
  )
tab_08_03
```

```{r}
library(flextable)
tab_08_03_ft <- tab_08_03 |>
  subset(select = -diabcd_f) |>
  flextable() |>
  autofit() |>
  add_header_lines(wrap_long_lines(
    "Analysis Set: Full Analysis Set - Observed Cases at baseline"
  )) |>
  set_caption(
    caption = wrap_long_lines(
      "Tab 08 03 - Analysis of Variance / Least Means Square estimations
      (95% Confidence Interval) of Demographics Parameters and Baseline
      Anthropometrics by Diabetes Group for Supplementary Parameters"
    )
  ) |>
  footnote(
    part = "header",
    i = 2, j = 2,
    value = as_paragraph(
      "Note: pval, p value of diabetes group effect test by F test."
    ),
    ref_symbols = "a"
  ) |>
  add_footer_lines("Source: ADSL and ADVS/ADLB observed cases at baseline.") |>
  theme_booktabs()
tab_08_03_ft
```

```{r}
bnm <- "tab_08_03"
dir_tab <- params$paths$tab
dir_dta <- params$paths$dta

file.path(dir_tab, paste0(bnm, "_ft.RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tab_08_03_ft, file = .)

file.path(dir_dta, paste0(bnm, ".RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tab_08_03, file = .)

file.path(dir_tab, paste(bnm, sep = ".", "docx")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_docx(tab_08_03_ft, path = .)

file.path(dir_tab, paste(bnm, sep = ".", "html")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_html(tab_08_03_ft, path = .)

file.path(dir_dta, paste(bnm, sep = ".", "csv")) %T>%
  message("[output] Table saved as ", .) %>%
  write.csv(tab_08_03, file = ., row.names = FALSE)
```


## Tab 08 04 - Post-hoc: Demographics and Baseline Anthropometrics by Diabetes Group (additional parameters)

```{r}
tab_08_04_raw <- ads |>
  subset(
    ! paramcd %in% c(
      "age", "trainn", "GLU0", "GLU30", "GLU60", "GLU120",
      "INSULIN0", "INSULIN30", "INSULIN60", "INSULIN120",
      "HBA1C", "HOMAB", "HOMAIR", "MATSUDA",
      "TRIG", "CHOL", "LDL", "HDL", "VO2MAXE", "WEIGHT", "BMI",
      "FATMASS", "BODYFATP", "LEANMASS", "VAT"
    )
  ) |>
  (\(x) split(x, f = x$paramcd, drop = TRUE))() |>
  lm_by_paramcd(indep_var = "diabcd") |>
  lsm_pairs_by_param(indep_var = "diabcd") |>
  (\(x) Reduce(rbind, x))()

library(tidyr)
tab_08_04 <- tab_08_04_raw |>
  subset(select = c(
    paramcd, contrast, estimate, SE, df, p.value, lower.CL, upper.CL
  ))

library(flextable)
tab_08_04_ft <- tab_08_04 |>
  (\(x) split(x, f = x$paramcd))() |>
  lapply(
    \(x) {
      x$p.value <- format(round(x$p.value, 5))
      x$estimate <- format(signif(x$estimate, 5))
      x$SE <- format(signif(x$SE, 6))
      x$lower.CL <- format(signif(x$lower.CL, 5))
      x$upper.CL <- format(signif(x$upper.CL, 5))
      x
    }) |>
  (\(x) Reduce(rbind, x))()|>
  flextable() |>
  fontsize(size = 9, part = "all") |>
  autofit() |>
  add_header_lines(
    "Analysis Set: Full Analysis Set - Observed Cases at baseline"
  ) |>
  set_caption(
    caption = wrap_long_lines(
      "Tab 08 04 - Post-hoc tests for the Analysis of Variance of
      Demographics and Baseline Anthropometrics by Diabetes Group"
    )
  ) |>
  footnote(
    value = as_paragraph(wrap_long_lines(
      "CL, 95% Confidence Limit; SE, Standard Error."
    )),
    ref_symbols = ""
  ) |>
  footnote(
    value = as_paragraph(wrap_long_lines(
      "Note: P value adjustment by Tukey's method for comparing a family of
      3 estimates."
    )),
    ref_symbols = ""
  )|>
  footnote(
    value = as_paragraph(wrap_long_lines(
      "Source: ADSL and ADVS/ADLB observed cases at
      baseline."
    )),
    ref_symbols = ""
  ) |>
  theme_booktabs()

tab_08_04_ft
```

```{r}
bnm <- "tab_08_04"
dir_tab <- params$paths$tab
dir_dta <- params$paths$dta

file.path(dir_tab, paste0(bnm, "_ft.RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tab_08_04_ft, file = .)

file.path(dir_dta, paste0(bnm, ".RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tab_08_04, file = .)

file.path(dir_tab, paste(bnm, sep = ".", "docx")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_docx(tab_08_04_ft, path = .)

file.path(dir_tab, paste(bnm, sep = ".", "html")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_html(tab_08_04_ft, path = .)

file.path(dir_dta, paste(bnm, sep = ".", "csv")) %T>%
  message("[output] Table saved as ", .) %>%
  write.csv(tab_08_04, file = ., row.names = FALSE)
```

## Session Informations

```{r}
sessioninfo::session_info()
```

