---
title: "micro-RNA: demographics  - deprecated"
author: "Francois Collin"
date: "`r format(Sys.time(), '%Y-%m-%d at %H:%M:%S (%Z)')`"
bibliography: references.bib
---

## Demographics:

Subject level variables, such as the age and number of training relied on
\index{analysis of variance}. The variables associated with repeated
measurements (measurement at baseline and after exercise) such as BMI and
diet were analysed in a mixed model to estimate the effect of the diabetes
status (D) and the exercise (E) while accounting for the random effect
associated with the patient; the interaction term DxE was investigated *via*
a log-likelihood ratio test. These models were used to provided least mean
square estimates so as to:

- verify the balanced panel of subjects.
- verify the leading differences between diabetes groups and due to the
  exercise.

```{r, settings}
params <- yaml::read_yaml("_prog.yml")
devtools::load_all("src/pkg/dbs.data")
devtools::load_all("src/pkg/latarnia.utils")
```

```{r, ads}
#| code-fold: show

adsl <- dbs.data::adsl
advs <- dbs.data::advs

adsl |> head() |> knitr::kable(caption = "Outlook ADSL")
advs |> head() |> knitr::kable(caption = "Outlook ADVS")
```

@Szczerbinski2022

## Table 02 01

```{r, t_02_01}
library(tidyr)
ads <- advs |>
  subset(
    dtype == "" & paramcd %in% c("BMI", "DCAL"),
    select = c(subjid, avisitn, avisit, paramcd, aval)
  ) |>
  pivot_wider(names_from = paramcd, values_from = aval) |>
  merge(adsl)

library(lme4)
mod_age <- lm(age ~ diabcd, data = adsl)
mod_train <- lm(trainn ~ diabcd, data = adsl)

mod_bmi_int <- lmer(BMI ~ diabcd * avisit + (1 | subjid), data = ads)
mod_bmi <- lmer(BMI ~ diabcd + avisit + (1 | subjid), data = ads)

mod_DCAL_int <- lmer(DCAL ~ diabcd * avisit + (1 | subjid), data = ads)
mod_DCAL <- res2 <- lmer(DCAL ~ diabcd + avisit + (1 | subjid), data = ads)

pval_format <- function(x, nm) as_tibble(x)
pval <- function(x, ...) UseMethod("pval", x)
pval.lm <- function(x, ...) {
  aov <- car::Anova(x)
  y <- t(aov[-nrow(aov), ncol(aov), drop = FALSE])
  pval_format(y)
}

pval.lmerMod <- function(x, lrt, ...) {
  aov <- car::Anova(x)
  y <- t(aov[, ncol(aov), drop = FALSE])
  # likelihood ratio test (‘p(DxE)‘)
  aov <- anova(x, lrt)
  y <- cbind(y, t(aov["lrt", "Pr(>Chisq)", drop = FALSE]))
  pval_format(y)
}

#' @examples
#' pval(mod_age)
#' pval(mod_train)
#' pval(mod_bmi, mod_bmi_int)
#' pval(mod_DCAL, mod_DCAL_int)

lms <- function(x, specs = ~diabcd) {
  eff_nm <- as.character(rev(specs)[[1]])
  y <- emmeans::emmeans(x, specs = specs)
  y <- as.data.frame(y)
  emm_nm <- c("emmean", paste(c("lower", "upper"), sep = ".", "CL"))
  y <- y[c(eff_nm[eff_nm %in% names(y)], emm_nm)]
  tidyr::pivot_wider(
    y,
    names_from = "diabcd",
    values_from = all_of(emm_nm)
  )
}

#' @examples
#' lms(mod_age, ~diabcd)
#' lms(mod_bmi, ~diabcd|avisit)

eff_pval <- function(x, specs, x2 = NULL) {
  eff <- lms(x, specs)
  p <- pval(x, x2)
  names(p) <- c(
    diabcd = "p (D)",
    avisit = "p (E)",
    lrt = "p (DxE)"
  )[names(p)]

  cbind(
    var = all.vars(terms(x))[[1]],
    dplyr::union_all(eff, p)
  )
}

nsmall <- function(x, nsmall)
  format(round(x, digits = nsmall), nsmall = nsmall)

format_mean_ci <- function(mean, lcl, ucl, digits = 1) {
  ifelse(
    is.na(mean),
    "",
    paste0(
      nsmall(mean, digits),
      " (", nsmall(lcl, digits), ", ", nsmall(ucl, digits), ")"
    )
  )
}

format_pval <- function(pval) {
  ifelse(is.na(pval), "", nsmall(pval, 3))
}

format_rows <- function(df, digits = 1)
  tibble(
    Variable = df$var,
    stat = df$stat,
    group = df$avisit,
    NGT = format_mean_ci(df$emmean_NGT, df$lower.CL_NGT, df$upper.CL_NGT, digits),
    IGT = format_mean_ci(df$emmean_IGT, df$lower.CL_IGT, df$upper.CL_IGT, digits),
    T2D = format_mean_ci(df$emmean_T2D, df$lower.CL_T2D, df$upper.CL_T2D, digits),
    `p (D)` = format_pval(df$`p (D)`),
    `p (E)` = format_pval(df$`p (E)`),
    `p (DxE)` = format_pval(df$`p (DxE)`)
  )

tbl <- Reduce(
  dplyr::union_all,
  list(
    eff_pval(mod_age, ~diabcd),
    eff_pval(mod_train, ~diabcd),
    eff_pval(mod_bmi, ~diabcd|avisit, mod_bmi_int),
    eff_pval(mod_DCAL, ~diabcd|avisit, mod_DCAL_int)
  )
) |>
  within(expr = {
    stat <- factor(
      is.na(emmean_NGT),
      levels = c("TRUE", "FALSE"),
      labels = c("Test", "Estimate")
    )
    avisit <- factor(
      avisit,
      levels = c(" ", "0_before", "1_after"),
      labels = c(" ", "Before exercise:", "After exercise:")
    )
    avisit[is.na(avisit)] <- " "
    var <- factor(
      var,
      levels = c("age", "trainn", "BMI", "DCAL"),
      labels = c("Age", "train", "BMI", "Diet")
    )
  }) |>
  (\(df) df[order(df$var, df$avisit, is.na(df$`p (D)`)), ])() |>
  (\(df) split(df, f = df$var))() |>
  lapply(
    FUN = \(x) {
      switch(unique(as.character(x$var)),
             Diet = format_rows(x, digits = 0),
             format_rows(x, digits = 1)
      )
    }) |>
  dplyr::bind_rows()

library(flextable)
flextable::init_flextable_defaults()

tbl_ft <- tbl |>
  within(Variable <- as.character(Variable))|>
  dplyr::bind_rows() |>
  as_grouped_data(groups = "Variable") |>
  flextable() |>
  autofit() |>
  footnote(
    j = 7, part = "header",
    value = as_paragraph(wrap_long_lines(
      "Differences among the groups (independently from the exercise)"
    )),
    ref_symbols = "a"
  ) |>
  footnote(
    j = 8, part = "header",
    value = as_paragraph(wrap_long_lines(
      "Differences linked to the exercise (independently from the group)"
    )),
    ref_symbols = "b"
  ) |>
  footnote(
    j = 9, part = "header",
    value = as_paragraph(wrap_long_lines(
      "Interaction tested by Likelyhood Ratio Test (LRT)."
    )),
    ref_symbols = "c"
  ) |>
  footnote(
    value = as_paragraph(wrap_long_lines(
      "Note: Least means square estimations acounts for diabcdetes effect for
      all variables. For variables BMI and Diet, estimations include the
      additive effect of the exercise; however, the LRT excluded any
      significant interaction, no interaction effect was accounted for in the
      estimations of BMI and Diet."
    )),
    ref_symbols = ""
  ) |>
  add_header_lines(wrap_long_lines(
      "Analysis Set: Full Analysis Set - Observed Cases"
    )) |>
  set_caption(
    caption = wrap_long_lines(
      "Tab 02 01 - Least Means Square estimations of Age, number of trainings,
    BMI and diet by diabetes group and exercise along with effect tests."
    )
  ) |>
  theme_booktabs()

tbl_ft

bnm <- "tab_02_01"
dir_tab <- params$paths$tab
dir_dta <- params$paths$dta

file.path(dir_tab, paste0(bnm, "_ft.RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tbl_ft, file = .)

file.path(dir_dta, paste0(bnm, ".RData")) %T>%
  message("[output] Table saved as ", .) %>%
  save(tbl, file = .)

file.path(dir_tab, paste(bnm, sep = ".", "docx")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_docx(tbl_ft, path = .)

file.path(dir_tab, paste(bnm, sep = ".", "html")) %T>%
  message("[output] Table saved as ", .) %>%
  save_as_html(tbl_ft, path = .)

file.path(dir_dta, paste(bnm, sep = ".", "csv")) %T>%
  message("[output] Table saved as ", .) %>%
  write.csv(tbl, file = ., row.names = FALSE)

```
